---
title: "Text Mining 2"
author: "Anthony Kamau"
toc: yes
output: pdf_document
---

---

```{r, message=FALSE}
# install.packages(c("text2vec"))
library(tidyverse)
library(tidytext)
library(text2vec)
```

# **Word Embedding**
To work with text, we need to transform it into numerical values. One way is to count
the words and weight their frequency (e.g. with TF-IDF). Another method is 
word embedding. Word embedding techniques such as word2vec and GloVe use 
neural networks to construct word vectors. With these techniques, words in similar
contexts are represented by similar numerical vectors.

## **Load data set**
We will use the Harry Potter books as our data. Installing the 
`harrypotter` package from github using `remotes`.
```{r}
# install.packages("remotes")
remotes::install_github("bradleyboehmke/harrypotter")

library(harrypotter) # Not to be confused with the CRAN palettes package
```

It is a data set with all Harry Potter books, containing the full texts of the 
first seven Harry Potter books, Each text is in a character vector with each 
element representing a single chapter. It is provided from the `harrypotter`
package written by `Bradley Boehmke`.

  - **Philosophers_stone**: Harry Potter and the Philosophers Stone, published 
  in 1997
  
  - **Chamber_of_secrets**: Harry Potter and the Chamber of Secrets, published in 
    1998
    
  - **Prisoner_of_azkaban**: Harry Potter and the Prisoner of Azkaban, published in 
    1999
    
  - **Goblet_of_fire**: Harry Potter and the Goblet of Fire, published in 2000
  
  - **Order_of_the_phoenix**: Harry Potter and the Order of the Phoenix, published 
    in 2003
    
  - **Half_blood_prince**: Harry Potter and the Half-Blood Prince, published in 2005
  
  - **Deathly_hallows**: Harry Potter and the Deathly Hallows, published in 2007
 

Load the first seven novels in the Harry Potter series  
```{r}

hp_books <- c("philosophers_stone", "chamber_of_secrets",
              "prisoner_of_azkaban", "goblet_of_fire",
              "order_of_the_phoenix", "half_blood_prince",
              "deathly_hallows")

hp_words <- list(
  philosophers_stone,
  chamber_of_secrets,
  prisoner_of_azkaban,
  goblet_of_fire,
  order_of_the_phoenix,
  half_blood_prince,
  deathly_hallows
) |>
  # names each list element
  set_names(hp_books) |>
  # converts each book to a data frame and merge into a single data frame
  map_df(as_tibble, .id = "book") |>
  # converts book to a factor
  mutate(book = factor(book, levels = hp_books)) |>
  # removes empty chapters
  filter(!is.na(value)) |>
  # creates a chapter id column
  group_by(book) |>
  mutate(chapter = row_number(book))

# head(hp_words)
```

## **Tokenize data frame**
We'll use `unnest_tokens` function from the `tidytext` package to tokenize the 
data frame
```{r}
hp_words <- hp_words |>
  unnest_tokens(word, value)

head(hp_words)
```
## **Remove stop words** 
We remove stop words from tokenized data frame 
```{r}
hp_words <- hp_words |> 
  # remove stop words
  anti_join(stop_words)

head(hp_words)
```
## **Vocabulary of unique terms**
Creating a vocabulary of unique terms using the `create_vocabulary` function from
the `text2vec` package and removing the words that appear less than 5 times.
```{r}
# make it a list (iterator)
hp_words_ls <- list(hp_words$word)

# create index-tokens
it <- itoken(hp_words_ls, progressbar = FALSE) 

# collects unique terms 
hp_vocab <- create_vocabulary(it)

# filters the infrequent terms (number of occurrence is less than 5)
hp_vocab <- prune_vocabulary(hp_vocab, term_count_min = 5)

# show the resulting vocabulary object (formatting it with datatable)
hp_vocab

# # We've just created word counts, that's all the vocabulary object is!
```

## **Token Co-occurrence Matrix (TCM)**
Creating a `token co-occurrence matrix (TCM)`. We first apply `vocab_vectorizer`
function to transform the list of tokens into a vector space. Then, use `create_tcm`
function to create a TCM with the window of 5 for context words.

```{r}
# maps words to indices
vectorizer <- vocab_vectorizer(hp_vocab)

# use window of 5 for context words
hp_tcm <- create_tcm(it, vectorizer, skip_grams_window = 5)


# Note that such a matrix will be extremely sparse. 
# Most words do not go with other words in the grand scheme of things. So when 
# they do, it usually matters.
```

## **Train word vectors**
Fit the word vectors on our data set, with embedding size (rank variable) equal 
to 50, and the maximum number of co-occurrences equal to 10. 
Train word vectors in 20 iterations. 

```{r}
glove      <- GlobalVectors$new(rank = 50, x_max = 10)
hp_wv_main <- glove$fit_transform(hp_tcm, n_iter = 20, convergence_tol = 0.001)
```
## **Combining and extracting word vectors**
The GloVe model learns two sets of word vectors: main and context. 
Essentially they are the same since the model is symmetric. In general combining 
the two sets of word vectors leads to higher quality embeddings. The best practice
is to combine both the main word vectors and the context word vectors into one matrix.
Extract the word vectors and save the summation of them.

```{r}
# extract context word vector
hp_wv_context <- glove$components

# check the dimension for both matrices
dim(hp_wv_main); dim(hp_wv_context)

# Either word-vectors matrices could work, but the developers of the technique
# suggest the sum/mean may work better

# transpose one matrix to perform matrix addition
hp_word_vectors <- hp_wv_main + t(hp_wv_context) 
```

## **Finding most similar words**
Find the most similar 10 words to each of the words: “harry”, “death”, and “love”

**HARRY**
```{r}
# extract the row of "harry"
harry <- hp_word_vectors["harry", , drop = FALSE]

# calculates pairwise similarities between"harry" and the rest of words
cos_sim_harry <- sim2(x = hp_word_vectors, y = harry, method = "cosine", norm = "l2")

# the top 10 words with the highest similarities
head(sort(cos_sim_harry[,1], decreasing = T), 10)
```

**DEATH**
```{r}
# extract the row of "death"
death <- hp_word_vectors["death", , drop = FALSE]

# calculates pairwise similarities between"harry" and the rest of words
cos_sim_death <- sim2(x = hp_word_vectors, y = death, method = "cosine", norm = "l2")

# the top 10 words with the highest similarities
head(sort(cos_sim_death[,1], decreasing = T), 10)
```

**LOVE**
```{r}
# extract the row of "love"
love <- hp_word_vectors["love", , drop = FALSE]

# calculates pairwise similarities between"harry" and the rest of words
cos_sim_love <- sim2(x = hp_word_vectors, y = love, method = "cosine", norm = "l2")

# top 10 words with the highest similarities
head(sort(cos_sim_love[,1], decreasing = T), 10)
```
## **More word vector manipulations**
Play around with the word vectors. For example, add the word vector of “harry” 
with the word vector of “love” and subtract them from the word vector of “death” 
and find the top terms in the result.

```{r}
# add/subtract word vectors
test <- harry + love - death

# calculates pairwise similarities between"harry" and the rest of words
cos_sim_test <- sim2(x = hp_word_vectors, y = test, method = "cosine", norm = "l2")

# top 10 words with the highest similarities
head(sort(cos_sim_test[,1], decreasing = T), 10)
```

# **Dictionary-based Sentiment Analysis**

We are going to use the data set `movie_review`. This is a data set with 5,000 
IMDB movie reviews available from the `text2vec` package, labeled according to
their IMDB rating. The sentiment of the reviews is binary, meaning an 
IMDB rating < 5 results in a sentiment score of 0, and a rating >=7 has a sentiment score of 1. 
No individual movie has more than 30 reviews.

## **Load dataset**
```{r}
data("movie_review")

movie_review <- as_tibble(movie_review)

head(movie_review)
```
## **Tidytext package 4 general purpose lexicons**
The `tidytext` package contains 4 general purpose lexicons in the sentiments dataset.

  - **afinn**: list of English words rated for valence between -5 and +5
  
  - **bing**: list of positive and negative sentiment
  
  - **nrc**: list of English words and their associations with 8 emotions (anger, fear, anticipation, trust, surprise, sadness, joy, and disgust) and 2 sentiments (negative and positive); binary
  
  - **loughran**: list of sentiment words for accounting and finance by category (Negative, Positive, Uncertainty, Litigious, Strong Modal, Weak Modal, Constraining)
  
We are going to use *labMT dictionary* (Dodds' et al. 2011), one of the best dictionaries for sentiment analysis (see e.g. \textcolor{blue}{\href{https://link.springer.com/content/pdf/10.1140/epjds/s13688-017-0121-9.pdf}{this paper}}.)


## **Download the labMT dictionary**
```{r, message=FALSE}
labMT_url <- "https://github.com/andyreagan/sentidict/raw/refs/heads/main/data/LabMT/labMT2english.txt.gz"

labMT <- readr::read_delim(labMT_url, delim = "\t") |> select(word, value = happs)
```
## **Tokenization**
Use `unnest_tokens` function from `tidytext` package to break the `movie_review`
data set into individual tokens.
```{r}
# tokenize the reviews
tidy_review <- 
  movie_review |> 
  unnest_tokens(word, review) 

head(tidy_review)
```
## **Stop words removal**
Remove the stop words from the tokenized data frame
```{r}
tidy_review <- tidy_review |> 
  # remove stop words
  anti_join(stop_words)

head(tidy_review)
```
## **Finding sentiment scores**
Using the `inner_join` function to find a sentiment score for each of the tokenized
review words using the `labMT dictionary`
```{r}
review_sentiment <- 
  tidy_review |>
  inner_join(labMT)

head(review_sentiment)
```
## **Average sentiment score for each review**
Calculating the average sentiment score for each review. Comment on the three 
most positive and negative reviews (i.e., which has the highest and lowest average
sentiment score)? 
```{r}
sorted_review_sentiment <- 
  review_sentiment |>
  # group by review id
  group_by(id) |> 
  # compute the average sentiment score
  summarize(average_sentiment = mean(value),
            sentiment = mean(sentiment)) |> 
  # arrange the average sentiment score in descending order
  arrange(desc(average_sentiment)) 

sorted_review_sentiment
```
**HAPPIEST REVIEWS**
```{r}
# filter happiest reviews
movie_review |> 
  filter(id %in% head(sorted_review_sentiment$id, 3)) |> 
  pull(review)
```
**SADDEST REVIEWS**
```{r}
# filter saddest reviews
movie_review |> 
  filter(id %in% tail(sorted_review_sentiment$id, 3)) |> 
  pull(review)
```

Bar chart of these average sentiment scores across the `ids`. Color used to show
the original sentiment.
```{r}
sorted_review_sentiment |>
  # create the coloring variable
  mutate(sentiment = ifelse(sentiment > 0, "pos", "neg")) |>
  # reorder `id` such that average sentiment scores are presented in descending order. 
  ggplot(aes(
    x = reorder(id, -average_sentiment), 
    y = average_sentiment - mean(average_sentiment), 
    fill = sentiment
  )) +
  # plot the bar chart using `geom_col`
  geom_col() + 
  labs(x = "id", title = "Average sentiment scores") +
  theme_minimal() + 
  theme(axis.text.x=element_blank())
```
Creating a predicted_sentiment column, such that_average score higher than 
5.75 is positive = 1 and average score lower than or equal to 5.75 is negative = 0. 
Then, using a confusion matrix to compare this predicted_sentiment to the 
original sentiment.
```{r}
review_sentiment_agg <- 
  sorted_review_sentiment |>
  # Dichotomize average sentiment score to match with the original sentiment scores
  # 1: scores higher than zero; 0: scores lower than or equal to zero
  mutate(predicted_sentiment = ifelse(average_sentiment > 5.75, 1, 0),)

## not too bad
table(
  true = review_sentiment_agg$sentiment, 
  predicted = review_sentiment_agg$predicted_sentiment
)

# Accuracy = 1653+1729/5000 = 67.64%
```
Following the same steps, but only for the reviews in the top or bottom 25% of 
the distribution of average sentiment (there are likely many neutral reviews).
Accessing our accuracy and whether our results improve.
```{r}
summary(sorted_review_sentiment$average_sentiment)
```


```{r}
review_sentiment_agg <- sorted_review_sentiment |>
  filter((average_sentiment > 5.945) | (average_sentiment < 5.528)) |> 
  # Dichotomize average sentiment score to match with the original sentiment scores
  # 1: scores higher than zero; 0: scores lower than or equal to zero
  mutate(predicted_sentiment = ifelse(average_sentiment > 5.75, 1, 0),)

## not too bad
table(
  true = review_sentiment_agg$sentiment, 
  predicted = review_sentiment_agg$predicted_sentiment
)

# Accuracy = (935+990)/2501 = 77.0%
# Our accuracy increases considerably
```

